---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
## The Analysis Of Retail Store Data Set-Cardio Fitness Project
## By Chinedu H Obetta of Great Learning Academy

***Import and Read the Dataset***

##Setting the working directory and importing dataset
```{r}
setwd("C:/Users/CHINEDU/Documents/GREAT LEARNING-UNIVERSITY OF TEXAS")
library(readr)
Cardio_Good_Fitness = read.csv("CardioGoodFitness.csv")
```

##Variable Identification such as the dimension, the names of the columns, the tail and head of the data set and other basic summaries of the data set. 

```{r}
dim(Cardio_Good_Fitness)
head(Cardio_Good_Fitness, 10)
tail(Cardio_Good_Fitness, 10)
names(Cardio_Good_Fitness)
```
##Understanding the structure of the dataset
```{r}
str(Cardio_Good_Fitness)
summary(Cardio_Good_Fitness)
```

##Analysis of the products. This requires the loading of ggplot2.
```{r}
library(ggplot2)
class(Cardio_Good_Fitness$Product)
table(Cardio_Good_Fitness$Product)
```

```{r}
ggplot(Cardio_Good_Fitness) +
  aes(x = Product) +
  geom_bar(position = "dodge", fill = "Blue") +
  labs(x = "Product Type", y = "Product Count", title = "Univariate Analysis-Cardio Fitness", subtitle = "Product Distribution (3.3a)") +
  theme_minimal()
```
## THE ANALYSIS OF THE CUSTOMER'S AGE
```{R}
boxplot(Cardio_Good_Fitness$Age)  #### there exists three outliers
hist(Cardio_Good_Fitness$Age)
```
#ADJUSTING THE OUTLIERS
```{r}
AGE<- subset(Cardio_Good_Fitness$Age, Cardio_Good_Fitness$Age< 40)
summary(AGE)
hist(AGE)
```

### CUSTOMER'S YEARS OF EDUCATION
```{r}
class(Cardio_Good_Fitness$Education)
unique(Cardio_Good_Fitness$Education) ## There are 8 unique objects in Education, hence, it should be classified as a factor. 

Cardio_Good_Fitness$Education <- as.factor(Cardio_Good_Fitness$Education)
ggplot(Cardio_Good_Fitness) +
  aes(x = Education) +
  geom_bar(fill = "#0c4c8a") +
  labs(title = "Univariate Analysis-Cardio Fitness", subtitle = "Years Of Education (3.3c)", caption = "Years Of Education") +
  theme_minimal()
```

### Taking a look at Marital Status
```{r}
summary(Cardio_Good_Fitness$MaritalStatus)
plot(Cardio_Good_Fitness$MaritalStatus)
ggplot(Cardio_Good_Fitness) +
  aes(x = MaritalStatus) +
  geom_bar(fill = "#0c4c8a") +
  labs(title = "Univariate Analysis-Cardio Fitness", subtitle = "Distribution of Marital Status-3.3d", caption = "Distribution of Marital Status") +
  theme_minimal()
```

## ANALYSIS OF THE WEEKLY USAGE
```{r}
unique(Cardio_Good_Fitness$Usage) ## There are 6 unique categories of usage out of the 180 observations, hence, it is a factor variable.
Cardio_Good_Fitness$Usage <- as.factor(Cardio_Good_Fitness$Usage)

ggplot(Cardio_Good_Fitness) +
  aes(x = Usage) +
  geom_bar(fill = "#0c4c8a") +
  labs(title = "Univariate Analysis-Cardio Fitness", subtitle = "Weekly Usage-3.3e") +
  theme_minimal()
```
## ANALYSIS OF THE CUSTOMER'S INCOME AND MILES
```{r}
hist(Cardio_Good_Fitness$Income)
hist(Cardio_Good_Fitness$Miles)
boxplot(Cardio_Good_Fitness$Miles) # to review potential outliers.
#Removing Outliers
Miles <- subset(Cardio_Good_Fitness$Miles, Cardio_Good_Fitness$Miles< 120)
boxplot(Miles)
hist(Miles)
```
## ANALYSIS AND VISUAL DISPLAY OF GENDER
```{r}
table(Cardio_Good_Fitness$Gender)
plot(Cardio_Good_Fitness$Gender)
```
## GRAPHICAL ANALYSIS CUSTOMER'S FITNESS
```{r}
unique(Cardio_Good_Fitness$Fitness) ### There are 5 categories or levels of fitness, thus, "Fitness" is a factor variable.
Cardio_Good_Fitness$Fitness <- as.factor(Cardio_Good_Fitness$Fitness)
ggplot(Cardio_Good_Fitness) +
  aes(x = Fitness) +
  geom_bar(fill = "#0c4c8a") +
  labs(x = "Rating", title = "Univariate Analysis-Cardio Fitness", subtitle = "Fitness - Self Rated Fitness Score") +
  theme_minimal()
```
***BIVARIATE ANALYSIS***
## THE GRAPHICAL ANALYSIS OF WEEKLY USAGE.OF.MILES_vs_GENDER
```{r}
ggplot(Cardio_Good_Fitness) +
  aes(x = Usage) +
  geom_bar(fill = "Red") +
  labs(title = "MILE USAGE_VS_GENDER", caption = "Fig b1") +
  theme_minimal() +
  facet_wrap(vars(Gender))
```
##THE CHOICE OF THE PRODUCTS BY GENDER
```{r}
ggplot(Cardio_Good_Fitness) +
  aes(x = Product, fill = Gender) +
  geom_bar() +
  scale_fill_hue() +
  labs(title = "THE CHOICE OF THE PRODUCT TYPE BY GENDER", caption = "Fig b2") +
  theme_minimal() +
  facet_wrap(vars(Gender))
```
##THE EFFECT OF CUSTOMER’S LEVEL OF INCOME 
```{r}
ggplot(Cardio_Good_Fitness) +
  aes(x = Product, y = Income) +
  geom_boxplot(fill = "#0c4c8a") +
  labs(x = "Product Type", y = "Income Level", caption = "THE EFFECT OF CUSTOMER’S LEVEL OF INCOME ") +
  theme_minimal()
```
##THE CUSTOMER’S STATE OF FITNESS
```{r}
ggplot(Cardio_Good_Fitness) +
  aes(x = Product, fill = Fitness) +
  geom_bar() +
  scale_fill_hue() +
  labs(x = "Product Type") +
  theme_minimal() +
  facet_wrap(vars(Fitness))
```
## USAGE BY PRODUCT-----
```{r}
ggplot(Cardio_Good_Fitness, aes(x= Usage, fill = Product)) +
  geom_bar( position = "dodge") +
  labs(title = "Usage By Products")
names(Cardio_Good_Fitness)
```
## Understanding the relationship between income and gender
```{r}
ggplot(Cardio_Good_Fitness, aes(x=Gender, y= Income, color=MaritalStatus)) +
  geom_boxplot(outlier.shape = NA)+      #### assuming outliers are negligbles
  labs(title = "Relationship Between Income and Gender")
````
